---
output: github_document
---
<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
options(cli.hyperlink = FALSE)
```

# fastcpd

[![CRAN status](https://www.r-pkg.org/badges/version/fastcpd)](https://cran.r-project.org/package=fastcpd)
[![R-CMD-check](https://github.com/doccstat/fastcpd/workflows/R-CMD-check/badge.svg)](https://github.com/doccstat/fastcpd/actions)
[![Codecov test coverage](https://codecov.io/gh/doccstat/fastcpd/branch/main/graph/badge.svg)](https://app.codecov.io/gh/doccstat/fastcpd?branch=main)
[![Last Commit](https://img.shields.io/github/last-commit/doccstat/fastcpd)](https://github.com/doccstat/fastcpd)
[![r-universe](https://doccstat.r-universe.dev/badges/fastcpd)](https://doccstat.r-universe.dev)
[![CodeFactor](https://www.codefactor.io/repository/github/doccstat/fastcpd/badge)](https://www.codefactor.io/repository/github/doccstat/fastcpd)

## Overview

The fastcpd (**fast** **c**hange **p**oint **d**etection) is a fast implmentation
of change point detection methods in R. The __fastcpd__ package is designed to
find change points in a fast manner. It is easy to install and extensible to
all kinds of change point problems with a user specified cost function apart
from the built-in cost functions.

If you'd like to learn how to use the fastcpd effectively, please refer to:

* [Sequential Gradient Descent and Quasi-Newton's Method for Change-Point Analysis](https://proceedings.mlr.press/v206/zhang23b.html)

[documentation & examples](https://fastcpd.xingchi.li/reference/fastcpd.html#ref-examples)

## Installation

```{r eval = FALSE}
# Install from CRAN, (not yet available)
install.packages("fastcpd")
```

```{r eval = FALSE}
# Install the development version from GitHub using either of the following
# commands:

## install.packages("pak")
pak::pak("doccstat/fastcpd")

## install.packages("devtools")
devtools::install_github("doccstat/fastcpd")
```

## Dependency

`library(fastcpd)` depends on the following packages:

* [Rcpp](https://github.com/RcppCore/Rcpp), for C++ source code compilation.
* [RcppArmadillo](https://github.com/RcppCore/RcppArmadillo), for fast linear algebra.
* [fastglm](https://github.com/jaredhuling/fastglm), for fast generalized linear models.
* [DescTools](https://github.com/AndriSignorell/DescTools), for Winsorizing Poisson data.
* [glmnet](https://glmnet.stanford.edu/), for penalized regression.
* [ggplot2](https://github.com/tidyverse/ggplot2), for data visualization.

If you're compiling from source, you can run the following command to see the
complete set of system packages needed on your machine.

```{r eval = TRUE}
pak::pkg_sysreqs("doccstat/fastcpd")
```

## Contact us

Encountered a bug or unintended behavior?

1.  File a ticket at [GitHub Issues](https://github.com/doccstat/fastcpd/issues).
2.  Contact the developers specified in [DESCRIPTION](/DESCRIPTION).
