---
title: "Custom logistic regression model"
output: rmarkdown::html_vignette
description: |
  Examples of using custom model to reproduce change point detection in
  logistic regression models.
vignette: >
  %\VignetteIndexEntry{Custom logistic regression model}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(fastcpd)
```

```{r package_installation}
for (package in c("ggplot2", "mvtnorm")) {
  if (!requireNamespace(package, quietly = TRUE)) utils::install.packages(
    package, repos = "https://cloud.r-project.org", quiet = TRUE
  )
}
```

We try to reproduce the logistic regression models with
custom cost functions and show that the results are similar to the built-in
logistic regression models.

The built-in logistic regression model in `fastcpd()` is implemented with the
help of the **fastcpd** package **fastglm**. The **fastcpd** package
utilizes the iteratively reweighted least squares with the step-halving approach
to help safeguard against convergence issues. If a
custom cost function is used with gradient descent, we should expect the results
may not be the exact same as the built-in logistic regression model.

Specifying the `cost`, `cost_gradient` and `cost_hessian`
parameters below,
we can obtain similar results as the built-in logistic regression
model.

```{r fastcpd_binomial}
set.seed(1)
x <- matrix(rnorm(1500, 0, 1), ncol = 5)
theta <- rbind(rnorm(5, 0, 1), rnorm(5, 2, 1))
y <- c(
  rbinom(125, 1, 1 / (1 + exp(-x[1:125, ] %*% theta[1, ]))),
  rbinom(175, 1, 1 / (1 + exp(-x[126:300, ] %*% theta[2, ])))
)
binomial_data <- data.frame(y = y, x = x)

result <- fastcpd.binomial(cbind(y, x), r.progress = FALSE, cost_adjustment = NULL)
summary(result)
```

```{r custom_cost_functions}
logistic_loss <- function(data, theta) {
  x <- data[, -1]
  y <- data[, 1]
  u <- x %*% theta
  nll <- -y * u + log(1 + exp(u))
  nll[u > 10] <- -y[u > 10] * u[u > 10] + u[u > 10]
  sum(nll)
}
logistic_gradient <- function(data, theta) {
  x <- data[nrow(data), -1]
  y <- data[nrow(data), 1]
  c(-(y - 1 / (1 + exp(-x %*% theta)))) * x
}
logistic_hessian <- function(data, theta) {
  x <- data[nrow(data), -1]
  prob <- 1 / (1 + exp(-x %*% theta))
  (x %o% x) * c((1 - prob) * prob)
}
result <- fastcpd(
  y ~ . - 1, binomial_data, epsilon = 1e-5, cost = logistic_loss,
  cost_gradient = logistic_gradient, cost_hessian = logistic_hessian,
  r.progress = FALSE
)
summary(result)
```

Note that the result obtained through custom cost functions is inferior
compared to
the one obtained through built-in models.
We remark that the results can be improved with extra parameters
already provided in the package.
The detailed discussion of several advanced usages of the package can be found
in
[Advanced examples](https://fastcpd.xingchi.li/articles/examples-advanced.html).

# Appendix: all code snippets

```{r ref.label = knitr::all_labels(), echo = TRUE, eval = FALSE}
```
